---
title: "final_mock"
format: html
---

## problem 1
In this question we will use the iris dataset (included in R) to perform k-means clustering. The following code should be included to first process the data (as we will only use 2
predictors).
```{r}
## write answer here
data(iris)
iris_data <- iris[, c(1, 2)]

```


Problem 1a
Using iris_data, perform k-means clustering with 𝑘 = 3 clusters in the data. Be sure to set
the maximum number of iterations to 100 and perform at least 10 random starts. Save the
estimated clusters as iris_kmeans.
```{r}
## write answer here
iris_kmeans <- kmeans(iris_data, centers = 3, iter.max = 100, nstart = 10)
```


Problem 1b
Plot the data in iris_data as a scatter plot, where the points are colored by the estimated
clusters you obtain from iris_kmeans.

```{r}
## write answer here
plot(iris_data$Sepal.Length, iris_data$Sepal.Width, col = iris_kmeans$cluster)
```


Problem 2
The stringr package contains a dataset sentences consisting of some short sentences. We
will use a subset of these words, which can be obtained with the following code.
```{r}
## write answer here
library(stringr)
small_sentences <- sentences[1:10]
words_vector <- unlist(str_split(small_sentences, "\\s+"))[1:20]
```


Problem 2a
Using words_vector, construct a regular expression to extract the first 4 letter word from
each of these sentences. You can show a vector where the element is either a 4 letter words in
words_vector or NA if the corresponding element in words_vector is not a 4 letter word.
```{r}
## write answer here
str_extract(words_vector, pattern = "^....$")
```


Problem 2b
Again using words_vector, return all words which start with the letter b or c
```{r}
## write answer here
words_vector[str_detect(words_vector, pattern = "^[b-c]")]
```


Problem 2c
Using words_vector, extract a vector containing only words which have additional characters
beyond the 26 letters in the alphabet.
```{r}
## write answer here
words_vector[str_detect(words_vector, pattern = "\\W")]
```



Suppose you were given a statistical model which had the following log-likelihood function,
log ℒ = −𝑛𝜃 + log(𝜃)
𝑛
∑
𝑖=1
𝑥𝑖
,
where 𝜃 is the parameter in this model, 𝑛 is the number of observations and each of these
observations are 𝑥1
, 𝑥2
, … , 𝑥𝑛.
Here log is the natural logarithm.

Problem 3a
Write a function pois_llh which will take in as input values of theta,n, and a vector x and
compute the log-likelihood. What does this return for the values pois_llh(theta = 1, n =
5, x = c(1, 2, 3, 4, 5))?
```{r}
## write answer here
pois_llh <- function(theta, n, x) {
llh <- -n * theta + log(theta) * sum(x)
llh
}
pois_llh(theta = 1, n = 5, x = c(1:5))
```



Problem 3b
With n and x fixed at the values in the previous part, evaluate this function at the range of
values of theta_vals = seq(from = 0.1, to = 8, length.out = 100). Construct a plot
of these values of the log-likelihood and identify the maximum likelihood estimate using the
plot. You do not have to identify it exactly, just an educated guess based on the plot.
```{r}
## write answer here
theta_vals <- seq(from = 0.1, to = 8, length.out = 100)
lik_vals <- pois_llh(theta_vals, n = 5, x = 1:5)
plot(theta_vals, lik_vals, type = "l")
```

Problem 4
In this problem we will use KNN to do classification, predicting a binary variable purchased
in ad_data.csv (download instructions are given on the first page).
The following code will split this data into training and test data, assuming you have read it
in as ad_data.
```{r}
## write answer here
ad_data <- read.csv("ad_data.csv")
set.seed(100)
prop_train <- 0.7
prop_test <- 1 - prop_train
n_data <- nrow(ad_data)
train_test <- sample(c("Train", "Test"),
size = n_data, replace = TRUE,
prob = c(prop_train, prop_test))
training_data <- ad_data[train_test == "Train", ]
testing_data <- ad_data[train_test == "Test", ]
```


Problem 4a
Fit KNN on this training data where we use the variable EstimatedSalary as the single
preddictor, with 𝐾 = 5 and store the fit in an object called fit_knn. To do this you also need
to pass in the predictors in testing_data. (3 points)

```{r}
## write answer here
library(class)
fit_knn <- knn(train = as.matrix(training_data$EstimatedSalary),
test = as.matrix(testing_data$EstimatedSalary),
cl = training_data$Purchased, k = 5)
```


Compute the Accuracy of this KNN model on the test data. (2 points)
```{r}
## write answer here
mean(fit_knn != testing_data$Purchased)

```