---
title: "lab05"
format: html
---

Problem 1
This problem is about bird sightings in Canada. The files ebird30.csv and names.csv contain
data contributed by citizen scientists in 30 days to eBird: A resource maintained by the Cornell
Lab of Ornithology. The file ebird30.csv is a comma separated variable file with headers, and
one row per bird species. The first column is named Code and contains the code for each bird
species available in the data (these codes were created by eBird). The remaining columns are
named according to the provinces and territories (regions) of Canada (for example, CA.NU is
Nunavut). For each bird species and region, the corresponding entry of ebird30.csv indicates
the number of birds of that species observed in that region in the past 30 days. The file
names.csv is a comma separated variable file with headers, and one row per bird species. The
first column is the code for a species of bird, and the remaining columns are the common name
and scientific name for that bird. Included in the archive for this lab is an R script canada.R.
This script contains a function canada that will make a heatmap of Canada, showing values
for each province. This script requires that several packages be installed (they can be installed
using the script packages.R)
1
Problem 1a)
How many bird species have been spotted in only one region of Canada? Verify that your
number is correct by using the Canada function to plot the counts of one the bird species you
identify as being spotted in only one region of Canada (all regions should be dark blue except
one). Provide the plot.

```{r}
## write answer here
source("canada.R")
birds <- read.csv("ebird30.csv")
names <- read.csv("names.csv")
one_region <- which(apply(birds[, 2:14], 1, function(x) sum(x > 0)) == 1)
one_region

name <- names[one_region[1], ]
comm_name <- name$Common.Name
heat <- list()
for (i in 1:13) {
heat[[colnames(birds)[i+1]]] = birds[one_region[1], colnames(birds)[i+1]]
}
canada(title = comm_name, heat = heat)

```

Problem 1b)
Which bird species has the highest count, summed across all regions? Give the common name,
and the scientific name
```{r}
## write answer here
total_count <- apply(birds[, 2:14], 1, sum)
names[which.max(total_count), ]
```

Which bird species has the smallest non-zero count, summed across all regions? If there are
more than one, provide the number of bird species that are tied for smallest non-zero count,
and sort their common names in alphabetical order and list the first 10. (Hint: The sort
function applied to a vector of strings will return a sorted version of the strings in alphabetical
order.)

```{r}
## write answer here
total_count <- apply(birds[, 2:14], 1, sum)
smallest <- names[which(total_count == 1), ]$Common.Name
sort(smallest)[1:10]

```

