---
title: "Untitled"
format: html
---
Problem 1a
Looking only at the book “Sense and Sensibility”, determine the proportion of words in this
text which contain 10 or more letters. You can include all words (for this text) that are
contained in book_words. Note that we did not cover a function to do this in class, so you
should consult the documents for the stringr package available here.
```{r}
## write answer here
ss_book <- book_words[book_words$book == "Sense & Sensibility", ]
n_words <- nrow(ss_book)
n_words

ss_book$length <- str_length(ss_book$word)
sum(ss_book$length > 9)/n_words


```

Problem 1b
Now compute this quantity for each of the books in book_words. Similarly, determine the year
each book was published. Construct a plot where the x axis is the publication year and the y
axis is the proportion of words with 10 or more letters for each of the books in book_words.
(5 points)
Hint: There are many ways you could get the year a book was published. For some of these
books it may be contained in this dataset.

```{r}
## write answer here
prop <- book_words |>
group_by(book) |>
mutate(length = str_length(word)) |>
summarise(prop_long = sum(length > 9)/ n())
## can get some of this from the data
# some_years <- book_words |>
# group_by(book) |>
# mutate(line = row_number()) |>
# mutate(year = str_extract(word, pattern = "\\d{4}")) |>
2
# drop_na() |>
# filter(line < 8) |>
# select(book, year)
#
#
# others <- tibble(book = c("Emma", "Pride and Prejudice"),
# year = c("1815", "1813"))
#
# years <- bind_rows(some_years, others)
prop$year <- c(1811, 1813, 1814, 1815, 1803, 1818)
prop |>
ggplot(aes(year, prop_long)) +
geom_point() +
theme_bw()

```

Problem 1c
Now we will slightly modify the previous problem. For all books in this dataset, find all seven
letter words which start with the last 4 letters in the alphabet and end with e. How many
such unique words are there, and what are they?
```{r}
## write answer here
words <- book_words |>
group_by(book) |>
mutate(line = row_number()) |>
mutate(word = str_extract(word, pattern = "^[w-z].....e$")) |>
drop_na() |>
pull(word)
unique(words)
```

