---
title: "lab04"
format: html
---

Problem 1a)
Write an R function called my.lm that takes a vector x and a vector y and returns the slope a
and y-intercept b of the line 𝑓(𝑥) = 𝑎𝑥 + 𝑏 that minimizes the sum of squared errors between
𝑓(𝑥𝑖
) and $ y_i$. Your function must not make use of the R functions lm, lm.fit, or any R
function that does linear regression (you may, however, use functions like sum or mean). Your
function should return a named list with two elements named a and b. Provide your code.
You may use the following code as a template:

```{r}
## write answer here
my_lm <- function(x, y) {
mean_x <- mean(x)
mean_y <- mean(y)
b <- sum( (x - mean_x) * (y - mean_y) )/ sum ( (x - mean_x) ^2 )
a <- mean_y - b * mean_x
return(list(a = a, b = b))
}

```


Problem 1b)
Verify that the function my_lm that you wrote in Problem 1a) is correct by running my_lm on
a vector x of length three with values given by the first three digits of your student number
(one value per digit), and a vector y given by the last three digits of your student number.
Print the values that my_lm returns. Then, do the same regression with the R function lm and
print the resulting slope and y-intercept. Note whether or not the values returned by the two
methods are the same.
```{r}
## write answer here
x <- c(1, 2, 3)
y <- c(7, 8, 9)
my_lm(x, y)

lm(y ~ x)
```


Problem 2)
In the canvas folder for this lab is the SQLite database xcoretail.sqlite discussed in class.
For each month in the range January 2011 to August 2011 (inclusive), calculate Z Co Retail’s
total revenue for that month. Provide code, and plot the revenue by month (your y-axis can
be 1, 2, 3, … instead of dates, but make sure to label the axes clearly with units).
A description of the database follows. This database has two tables: sales and prices. The
table sales contains one row for each sale that the company made during the reporting period.
The table sales has three columns:
• StockCode: a string indicating the product identifier.
• Quantity: an integer indicating how much product was sold.
• InvoiceDate: a string indicating the date of the sale.
2
The table prices is a table indicating the price of each product in the company’s inventory
and has four columns:
• StockCode: a string indicating the product identifier (matching the identifiers in the
table sales).
• Description: a string describing the product.
• UnitPrice: the price in USD of the product.
While we want to eventually plot the total sales across all products, we will first do this for a
single product.

Part a)
Using the class code as a template, construct a plot showing the total sales of the item with
stock code 71053 for the 8 months from 2011-01 to 2011-08. To do this you will need to first
extract the total number of items sold each month and then, using information in the prices
table, compute the value of these sales, in USD. Plot this revenue against the month.

```{r}
## write answer here
library(RSQLite)
library(DBI)
## multiple ways to solve this
db <- dbConnect(SQLite(), dbname = "xcoretail.sqlite")
price_query <- "SELECT UnitPrice FROM prices WHERE StockCode = '71053'"
request <- dbSendQuery(db, price_query)
item_price <- as.numeric(dbFetch(request))
dbClearResult(request)
item_price

months <- paste0("2011-0", 1:8)
counts <- rep(NA, 8)
sales <- rep(NA, 8)
for(i in 1:8){
query1 <- sqlInterpolate(db, "SELECT SUM(Quantity) AS Total
FROM sales
3
WHERE InvoiceDate LIKE ?month
AND StockCode = '71053'",
month = paste0(months[i], "%"))
request <- dbSendQuery(db, query1)
counts[i] <- as.numeric(dbFetch(request))
sales[i] <- counts[i] * item_price
dbClearResult(request)
}
plot(x = 1:8, y = sales, type = "l", ylab = "Revenue (USD)", xlab = "Month")
```


Part b)
Now, potentially using a similar approach to Part a, compute the total revenue per month
across all items in this database.
```{r}
## write answer here
total_sales <- data.frame()
months <- paste0("2011-0", 1:8)
for(i in 1:8){
query1 <- sqlInterpolate(db, "SELECT StockCode, SUM(Quantity) AS Total
FROM sales
4
WHERE InvoiceDate LIKE ?month
GROUP BY StockCode",
month = paste0(months[i], "%"))
request <- dbSendQuery(db, query1)
curr_month <- dbFetch(request)
curr_month$Month <- months[i]
total_sales <- rbind(total_sales, curr_month)
dbClearResult(request)
}
total_sales$Sales <- NA
prices <- dbReadTable(db, "prices")
for(i in 1:nrow(total_sales)){
item_price <- prices[prices$StockCode == total_sales$StockCode[i],
"UnitPrice"]
total_sales$Sales[i] <- total_sales$Total[i] * item_price
}
month_sales <- data.frame(month = months, total_sales = NA)
for(i in 1:length(months)){
monthly_sales <- total_sales[total_sales$Month == months[i], "Sales"]
month_sales$total_sales[i] <- sum(monthly_sales)
}
plot(x = 1:8, y = month_sales$total_sales, type = "l", xlab = "Months",
ylab = "Revenue (USD)")
```


