---
title: "lab07"
format: html
---

Problem 1 (Clustering)
In class last week we wrote an R function my_dist, to compute the distance between points
and a set of centroids. This is a key component in the implementation of k-means clustering.
In this problem we will build on that code to implement k-means clustering from scratch.
Problem 1a
Solve the first Exercise problem in last weeks slides, where you write an R function my_dist
which will return the distance between 𝑛 data points and 𝐾 centroids in an 𝑛 × 𝐾 matrix.
```{r}
## write answer here
my_dist <- function(data, centroids) {
n_data <- nrow(data)
1
K <- nrow(centroids)
dist_mat <- matrix(NA, nrow = n_data, ncol = K)
for(i in 1:n_data){
for(k in 1:K) {
## compute distance between point i and centroid k
dist_mat[i, k] <- dist(rbind(data[i,], centroids[k, ]))
}
}
dist_mat
}
set.seed(100)
peng_scaled <- read_csv("peng_scaled.csv")


centroids <- sample_n(peng_scaled, size = 3)
head(my_dist(peng_scaled, centroids))

```


Problem 1b
Write your own implementation of the k-means without using any libraries and without using
the R function kmeans, as a function called my_kmeans. It should work for datasets of arbitrary
dimension D, and should return the centroids and the cluster assignments of the last iteration
of the algorithm. You may work from the code from the slides for Week 9 (but extend
2
from 2 dimensions to arbitrary dimension, and use your own my_dist function). For stopping
condition, have your function take two parameters (in addition to a parameter for the data):
• A maximum number of iterations such that if the number of iterations reaches this
maximum number, the iterations stop,
• A threshold such that if the all Euclidean distances between the centroids before an
iteration and after an iteration is less than this threshold, the iterations stop.
Set reasonable default values for these parameters. Provide your code, and write a one paragraph help for your function, indicating the names of the parameters, and the nature of the
return value.

```{r}
## write answer here
my_kmeans <- function(data, K = 3, max_iter = 100, threshold = 0.01) {
# Convert data to a matrix if it's not already
data <- as.matrix(data)
# Number of data points and dimension
n <- nrow(data)
d <- ncol(data)
# Randomly initialize centroids by sampling k points from data
set.seed(123) # You can remove or change the seed if you wish
init_indices <- sample(n, K)
centroids <- data[init_indices, , drop = FALSE]
# Vector to store cluster assignments
clusters <- rep(0, n)
# K-means iterations
for (iter in 1:max_iter) {
# Store old centroids for convergence check
old_centroids <- centroids
# 1) Assign each point to the cluster with the closest centroid
curr_dist <- my_dist(data, centroids)
clusters <- apply(curr_dist, 1, which.min)
# for (i in 1:n) {
# distances <- numeric(k)
# for (j in 1:k) {
# distances[j] <- my_dist(data[i, ], centroids[j, ])
# }
3
# clusters[i] <- which.min(distances)
# }
# 2) Update centroids based on current assignments
for (j in 1:K) {
cluster_points <- data[clusters == j, , drop = FALSE]
if (nrow(cluster_points) > 0) {
centroids[j, ] <- colMeans(cluster_points)
} else {
# If no points assigned to the cluster, keep the old centroid
centroids[j, ] <- old_centroids[j, ]
}
}
# 3) Check convergence (max distance between old & new centroids)
max_shift <- 0
for (j in 1:K) {
shift <- dist(rbind(old_centroids[j, ], centroids[j, ]))
if (shift > max_shift) {
max_shift <- shift
}
}
if (max_shift < threshold) {
break
}
}
# Return final assignments and centroids
return(list(
centroids = centroids,
clusters = clusters
))
}
```


Problem 1c
Now implement your clustering method for peng_scaled with 𝐾 = 3 clusters. Choose the
maximum number of iterations to be 100 and the threshold to be 0.001.
Compare the clusters from your implementation to those from the kmeans function, with the
corresponding values. Do they return the same clustering assignments? How can you tell?

```{r}
## write answer here
fit <- my_kmeans(data = peng_scaled, K = 3,
max_iter = 100, threshold = 0.001)
orig <- kmeans(peng_scaled, centers = 3, iter.max = 100, nstart = 10)
table(fit$clusters, orig$cluster)
```
