---
title: "lab06"
format: html
---

Problem 1
In this problem we will use data from the 1971 Canadian census, which looks at the relationship
between the average education and income of people in 100 different professions. This data is
included in the R package carData which you will need to install. Then, this data, in a data
frame Presitge, can be loaded in using the following code:
```{r}
## write answer here
library(carData)
data("Prestige")
```


Problem 1a
We will first split the data into a training and a test set, with ~75% in the training set and
~25% in the test set. Use set.seed(100). Write some code to do this, calling each the
training_set and the test_set in your R code.
1
```{r}
## write answer here
set.seed(100)
prop_train <- 0.75
prop_test <- 1 - prop_train
n_data <- nrow(Prestige)
train_test <- sample(c("Train", "Test"),
      size = n_data, replace = TRUE,
      prob = c(prop_train, prop_test))
training_set <- Prestige[train_test == "Train", ]
test_set <- Prestige[train_test == "Test",]

```


Problem 1b
Now fit a linear regression to this training data and compute the RMSPE when this regression
is used to predict on the test data. Be sure to print out this RMSPE.
```{r}
## write answer here
fit <- lm(income ~ education, data = training_set)
est_Y <- predict(fit, newdata = test_set)
RMSPE <- sqrt( mean( (est_Y - test_set$income) ^2) )
RMSPE

```

Problem 1c
Now we will tune KNN for this data, before later predicting for the test data. For K taking a
range of possible values from 𝐾 = 1 to 𝐾 = 50, use 5-fold CV to determine which value of K
gives the best CV-RMSPE on this data. Plot this result and clearly state the choice of 𝐾.

```{r}
## write answer here
library(caret)
2
Loading required package: ggplot2
Loading required package: lattice
K_vals <- seq(from = 1, to = 50, by = 1)
K_RMSPE <- rep(NA, length(K_vals))
CV_folds <- sample(1:5, size = nrow(training_set), replace = TRUE)
num_folds <- 5
for(j in 1:length(K_vals)) {
  k <- K_vals[j]
  CV_RMSPE <- rep(NA, num_folds)
    for(i in 1:num_folds){
      cv_train <- training_set[CV_folds != i, ]
      cv_val <- training_set[CV_folds == i, ]
      fit_knn <- knnreg(income ~ education, data = cv_train, k = k)
      cv_pred <- predict(fit_knn, cv_val)
      CV_RMSPE[i] <- sqrt( mean( (cv_pred - cv_val$income)^2) )
  }
  K_RMSPE[j] <- mean(CV_RMSPE)
}
plot(K_vals, K_RMSPE, type = "l")
which.min(K_RMSPE)
```

Problem 1d
For the best 𝐾 in the previous part, use this choice of 𝐾 to compute the RMSPE on the test
data. Does this model give better predictions than linear regression?
```{r}
## write answer here
fit_knn <- knnreg(income ~ education, data = training_set, k = 9)
est_Y_knn <- predict(fit_knn, test_set)
RMSPE_knn <- sqrt( mean( (est_Y_knn - test_set$income) ^2) )
RMSPE_knn
```