---
title: "lab03"
format: html
---

Problem 1: Hazardous waste
The material for this lab includes data about hazardous waste in Austria, Armenia and Azerbaijan acquired from data.un.org. In the sqlite file combined.sqlite a table called ‘lab03combined’
is provided. This table has 6 columns: ‘Country’, ‘Year’, ‘Generated’, ‘Incinerated’, ‘Landfilled’, ‘Recycled’. This table indicates the amount of hazardous waste generated, incinerated,
landfilled or recycled (in tonnes) in the corresponding year and country, according to the
UNSD/UNEP Questionnaire on Environment Statistics. For example, the first row of this
table is:
Country Year Generated Incinerated Landfilled Recycled
Armenia 1998 286,607 0 286,607 0
This row indicates that in 1998 in Armenia, 286,607 tonnes of hazardous waste were generated
according to the survey. The years provided in the sqlite file span 1995 to 2012. Note that not
all countries have the same set of provided years.
Part a)
As a first initial task you should take the data from incinerated.new, a csv file, and create a
new table from this data. Write the first 4 rows of this dataset to a new table in the database,
calling this table incin.


```{r}
## write answer here
library(RSQLite)
db <- dbConnect(SQLite(), dbname = "combined.sqlite")
incinerated <- read.csv("incinerated.new")
1
dbWriteTable(
conn = db,
name = "incin",
value = incinerated[1:4, ],
row.names = FALSE,
overwrite = TRUE
)

```


Part b)
Now use dbExecute and INSERT to add the final three rows of incinerated.new to this table
incin. Print out the final three rows of incin to confirm you have done this correctly.
```{r}
## write answer here
for(i in 5:7){
query <- paste0("INSERT INTO incin VALUES ('",
incinerated$Country[i], "', ", incinerated$Year[i],
", ", incinerated$Value[i],
"", ")")
dbExecute(db, query)
}
incin_data <- dbReadTable(db, "incin")
tail(incin_data, 3)
```


Part c)
Your next task is to update the original table in the database with all the new, more recent
data. The new data are provided in the comma separated variable files generated.new,
incinerated.new, landfilled.new, and recycled.new (corresponding to the respective
columns of the sqlite table). Write a R script that will update the sqlite table with these
new data by inserting the new data to the end of the table. One new row must be inserted
into the sqlite table for each unique Country/Year pair.
2
You can assume (initially) that each of these files contains the same number of
rows, and that each file will contain the same value of country and year in the
same rows.
For example, since the first row of data in generated.new is:
Armenia,2015,555075
and the first row of incinerated.new is:
Armenia,2015,3767
a new row must be inserted into the combined.sqlite lab03combined table with Country value
Armenia, Year value 2015, Generated value 555,075 and Incinerated value 3,767 (and so on
for Landfilled and Recycled, drawn from their corresponding .new files).
Write your code using dbExecute and INSERT statements. Do not hard code the INSERT
statements (i.e., if new .new files were provided again 5 years from now with the same format
and filenames, rerunning your code should further update the database to insert the next
version of the data, regardless of how many rows are in the new .new files.)
```{r}
## write answer here
db <- dbConnect(SQLite(), dbname = "combined.sqlite")
generated <- read.csv("generated.new")
landfill <- read.csv("landfilled.new")
recyled <- read.csv("recycled.new")
## since we can assume the number of rows is the same, each
## row in each file contains the same country and the same year
## we can just directly insert this data in row by row
for(i in 1:nrow(generated)){
query <- paste0("INSERT INTO lab03combined VALUES ('",
generated$Country[i], "', ", generated$Year[i],
", ", generated$Value[i],
", ", incinerated$Value[i],
", ", landfill$Value[i],
", ", recyled$Value[i],
"", ")")
dbExecute(db, query)
}
```


Part d)
Print the last 7 rows of the updated sqlite table. You can do this by dumping the updated
sqlite table into a R data frame, and then printing the last 7 rows of the data frame.
(2 points)

```{r}
## write answer here
updated_table <- dbReadTable(db, "lab03combined")
## note that even if you overwrite the data multiple times
## you will just keep adding the 7 rows at the end
## so you should get the same answer here regardless
tail(updated_table, 7)

```

```{r}
## write answer here

```