---
title: "midterm"
format: html
---
Dataset
For this exam we will make use of the following dataset, penguins.tsv, which you can download from the internet using the following commands.
download.file("https://owenward.github.io/files/penguins.tsv",
"penguins.tsv")
Note that this data may not exactly match your exam data, as the data has
changed.


Problem 1: Reading and Writing Data (6 points)
• Read penguins.tsv into your R session with the appropriate commands based on
the file format, and subset this data, keeping only the rows where the value of
flipper_length_mm is greater than 190. Store this subset data in an R object
penguin_sub. How many rows are in penguin_sub? (2 points)

```{r}
## write answer here
penguin_sub <- penguins[penguins$flipper_length_mm > 190, ]
nrow(penguin_sub)

```
Modify penguin_sub to remove the column bill_depth_mm. Store this data, without
this column, in an R object called penguin_mod. (2 points)
```{r}
## write answer here
penguin_mod <- penguin_sub[ ,-4]
```


Now, using penguin_mod, write this to a file penguins.psv, where the separator between
the columns in this dataset is given by .... (2 points)

```{r}
## write answer here
write.table(penguin_mod, file = "penguins.psv", sep = "...")

```



Problem 2: SQL Question (7 points)
• Using the original data in penguins.tsv, create an RSQL database, called peng.sqlite
and write the data in the tsv file into this database, calling in PENGUINS. (2 points)

```{r}
## write answer here
library(RSQLite)
db <- dbConnect(SQLite(), dbname = "peng.sqlite")
dbWriteTable(conn = db, name = "PENGUINS", value = penguins,
overwrite = TRUE)
```

Write a SQL command to select all rows from PENGUINS where flipper_length_mm is
greater than 190. Confirm that this query returns the same number of rows as the
corresponding R code you wrote in Problem 1. (3.5 points)

```{r}
## write answer here
request <- dbSendQuery(
db,
"SELECT * FROM PENGUINS WHERE flipper_length_mm > 190"
)
result <- dbFetch(request)
dbClearResult(request)
nrow(result)

```


Write a second SQL command to now group by the species variable and count the number
of rows and the average flipper length, across these groups. Print the output of this SQL
command. Make sure that your SQL command returns the species for each row in the
output. (1.5 points)

```{r}
## write answer here
request_avg <- dbSendQuery(
db,
"SELECT species, COUNT(*) AS n, AVG(flipper_length_mm) AS avg_flipp
FROM PENGUINS
GROUP BY species"
)

```


Problem 3: Linear Regression Question (7 points)
• Suppose we want to fit a simple linear regression model where we assume that there is
no intercept term. The regression model is then given by
𝑦𝑖 = 𝑓(𝑥𝑖
) + 𝜖𝑖
where 𝑓(𝑥𝑖
) = 𝛽𝑥𝑖
. The value of 𝛽 which minimizes the sum of squared error is given by
𝛽 =̂
∑
𝑛
𝑖=1 𝑥𝑖
𝑦𝑖
∑
𝑛
𝑖=1 𝑥
2
𝑖
.
Write a function beta_est which will compute this value of 𝛽 given vectors 𝑥 and 𝑦. What
value of beta does it return when your x variable is penguins$bill_length_mm and y is
penguins$flipper_length_mm.

```{r}
## write answer here
beta_est <- function(x, y) {
num <- sum(x * y)
den <- sum(x ^2)
beta <- num/den
beta
}
beta_est(penguins$bill_length_mm, penguins$flipper_length_mm)
```

For the model in the previous part (with only 𝛽), write a function to compute the sum
of squared errors and then use the optim function to determine the estimated value of
𝛽̂. Confirm this agrees with the result you got in the previous part (to 2 decimal places).
You can ignore any warnings the optim function gives you here and your initial value of
𝛽 for optim can be 0. (1 point)

```{r}
## write answer here
SSE <- function(par = 0, x, y) {
beta <- par
SSE <- sum( (y - beta * x)^2 )
SSE
}
est_fit <- optim(par = 0, SSE, x = penguins$bill_length_mm,
y = penguins$flipper_length_mm)
Warning in optim(par = 0, SSE, x = penguins$bill_length_mm, y = penguins$flipper_length_mm):use "Brent" or optimize() directly
est_fit$par

```

We have not considered linear regression with more than one predictor but the calculation
of the sum of squared error is largely similar. Suppose the regression model is
𝑦𝑖 = 𝛼 + 𝛽1𝑥𝑖 + 𝛽2
𝑧𝑖 + 𝜖𝑖
.
Write an R function, SSE_2, to compute the sum of squared errors for such a model. The
outline of this function is given below. Using y = penguins$flipper_length_mm, x =
penguins$bill_length_mm, z = penguins$bill_depth_mm, compute the SSE for this
function when alpha = 0, beta1 = 1, beta2 = 1.
```{r}
## write answer here
SSE_2 <- function(alpha, beta1, beta2, y, x, z) {
est <- alpha + beta1 * x + beta2 * z
sse <- sum( (y - est)^2)
sse
}
4
SSE_2(alpha = 0, beta1 = 1, beta2 = 1,
y = penguins$flipper_length_mm,
x = penguins$bill_length_mm,
z = penguins$bill_depth_mm)
```

