---
title: "Lab 07"
subtitle: "Due March 21, 5:00 PM PST"
author: "Brandon Cho - 301365816"
format: html
---


Modify this template file to create your submission.

# Problem 1a


```{r}
## write your answer here
my_dist <- function(data, centroids) {
  n <- nrow(data)
  K <- nrow(centroids)
  distances <- matrix(NA, n, K)
  for (j in 1:K) {

    distances[, j] <- sqrt(rowSums((data - matrix(centroids[j, ], n, ncol(data), byrow = TRUE))^2))
  }
  return(distances)
}
```


## Problem 1b
Parameters:
data -> the matrix of data to be clustered \
k -> number of clusters to create \
max_iter -> the maximum number of iterations the algorithm will run \
tol -> threshold such that if the all Euclidean distances between the centroids before an \
iteration and after an iteration is less than this threshold, the iterations stop \

Return value: \
A list of the centroid positions and a vector that explains which cluster a datapoint is assigned to

```{r}
## write answer here
my_kmeans <- function(data, K, max_iter, tol) {

  data <- as.matrix(data)
  n <- nrow(data)
  d <- ncol(data)
  max_iter <- 100
  tol <- 0.001
  
  initial_indices <- sample(1:n, K)
  centroids <- data[initial_indices, ]
  
  cluster_assignments <- rep(0, n)
  
  for (iter in 1:max_iter) {
    # compute distances
    distances <- my_dist(data, centroids)
    
    # assign clusters
    new_assignments <- apply(distances, 1, which.min)
    
    # compute mean for each cluster
    new_centroids <- matrix(NA, nrow = K, ncol = d)
    for (k in 1:K) {
      if (sum(new_assignments == k) == 0) {
        # If no points are assigned to cluster k, reinitialize that centroid randomly.
        new_centroids[k, ] <- data[sample(1:n, 1), ]
      } else {
        new_centroids[k, ] <- colMeans(data[new_assignments == k, , drop = FALSE])
      }
    }
    
    # compute shifts for each centroid
    centroid_shifts <- sqrt(rowSums((centroids - new_centroids)^2))
    
    # if shift is less the tol, stop loop
    if (all(centroid_shifts < tol)) {
      cat("Converged in", iter, "iterations.\n")
      centroids <- new_centroids
      cluster_assignments <- new_assignments
      break
    }
    
    centroids <- new_centroids
    cluster_assignments <- new_assignments
  }
  
  return(list(centroids = centroids, cluster_assignments = cluster_assignments))
}
```


## Problem 1c
The custom and built-in k_means functions may not always have identical results due to arbitrary clustering, however the contingency table shows the same points are joining the same cluster, showing that both the functions have similar results.


```{r}
## write answer here
peng_scaled <- read.csv("peng_scaled.csv")
data_matrix <- as.matrix(peng_scaled)

# run custom k-means.
set.seed(100)
custom_km <- my_kmeans(data_matrix, K = 3, max_iter = 100, tol = 0.001)

# run built in k-means
set.seed(100)
builtin_km <- kmeans(data_matrix, centers = 3, iter.max = 100)

# compare
comparison_table <- table(custom_km$cluster_assignments, builtin_km$cluster)
print(comparison_table)

```